```{r include=FALSE}
ALL <- read.table("/home/edisz/Documents/Uni/Projects/blog/quantitative_ecotoxicology/data/S33.csv", 
                 header = TRUE, 
                 sep = ";")
```

### Quantitative Ecotoxicology, page 33, example 2.1, Winsorization:

Get the data (Sulfate Concentrations from Savannah River (South Carolina) in mg / L)) from [here](https://raw.github.com/EDiLD/r-ed/master/quantitative_ecotoxicology/data/S33.csv) and read it into R:

```{r, eval=FALSE}
ALL <- read.table("S33.csv", 
                 header = TRUE, 
                 sep = ";")
```

So we have a data.frame with one variable and 21 observations:
```{r}
str(ALL)
ALL$SO4
```


Winsorization replaces extreme data values with less extreme values. I have written a small function to run the winsorisation:
```{r}
winsori <- function (x, width = 2)
{
  # check if sorted
  if(is.unsorted(x))
    stop("Values must be sorted!")
  # get number of observations
  n <- length(x)
  # Replace lowest
  x[1:width] <- x[width + 1]
  # replace highest
  x[(n - width + 1):n] <- x[(n-width)]
  x
}
```

The function takes a ordered vector and replaces the 2 highest and 2 lowest values (can be changed by the 'width'-Argument by their neighbors.

We can apply this function to our data and safe it as new column:
```{r}
ALL$SO4_win <- winsori(ALL$SO4)
# display the first and 5 last rows
ALL[c(1:5, 17:21), ]
```

Worked as expected.
The Winsorized mean and standard-deviation is:
```{r}
# mean
mean(ALL$SO4_win)
# standard deviation
sd(ALL$SO4_win)
```

For the Winsorized Standard Deviation we need again a homemade function:
```{r}
sw <- function(x, width = 2){
  n <- length(x)
  sd(x) * (n - 1) / (n - 2*width -1)
}
sw(ALL$SO4_win)
```

And lastly we calculate the mean for the trimmed data (remove two observation from each tail):
```{r}
mean(ALL$SO4, trim=2/21)
```

Code and data are available at my [github-repo](https://github.com/EDiLD/r-ed/tree/master/quantitative_ecotoxicology) under filename 'p33'.