### Quantitative Ecotoxicology, page 35, Robust Regression on Order Statistics:

Get the data from [here](https://raw.github.com/EDiLD/r-ed/master/quantitative_ecotoxicology/data/S35.csv) and read it into R:
```{r include=FALSE}
SO4 <- read.table("/home/edisz/Documents/Uni/Projects/blog/quantitative_ecotoxicology/data/S35.csv", 
                  header = TRUE, 
                  sep = ";")
```

```{r, eval=FALSE}
SO4 <- read.table("S35.csv", 
                 header = TRUE, 
                 sep = ";")
```


First we need to convert the vector indicating if an observation is censored to TRUE/FALSE:
```{r}
SO4$rem2 <- ifelse(SO4$rem == "<", TRUE, FALSE)
SO4
```

Then we can run the Robust Regression on Order Statistics with the ros() function from the NADA package:
```{r message=FALSE}
require(NADA)
rs <- ros(SO4$value, SO4$rem2)
print(rs)
```

Which gives the same mean and standard deviation as the SAS-Makro.

Code and data are available at my [github-repo](https://github.com/EDiLD/r-ed/tree/master/quantitative_ecotoxicology) under filename 'p35'.

