### Quantitative Ecotoxicology, page 39, example 2.3, Kaplan–Meier estimates

Get the data from [here](https://raw.github.com/EDiLD/r-ed/master/quantitative_ecotoxicology/data/p39.csv) and read it into R:
```{r include=FALSE}
SULFATE <- read.table("/home/edisz/Documents/Uni/Projects/blog/quantitative_ecotoxicology/data/p39.csv", 
                  header = TRUE, 
                  sep = ";")
```

```{r, eval=FALSE}
SULFATE <- read.table("p39.csv", 
                  header = TRUE, 
                  sep = ";")
```

Convert left to right censored data:
```{r}
SULFATE$FLIP <- abs(SULFATE$SO4 - 8)
SULFATE
```

The Kaplan-Meier estimates can be calculated using survfit() from the survival package:
```{r message=FALSE}
require(survival)
fit <- survfit(Surv(FLIP, FLAG) ~ 1, data = SULFATE, conf.type="plain") 
fit
```

I set conf.type="plain" to be concordant with 'CONFTYPE=LINEAR' from SAS.

The median of 3.15, 95% CI [1.8, 4.5] is the same as with SAS.

Finally a quick plot:
```{r p39}
plot(fit)
```


Code and data are available at my [github-repo](https://github.com/EDiLD/r-ed/tree/master/quantitative_ecotoxicology) under filename 'p39'.


