### Quantitative Ecotoxicology, page 42, example 2.4, Wilcoxon rank sum test

Get the data from [here](https://raw.github.com/EDiLD/r-ed/master/quantitative_ecotoxicology/data/p42.csv) and read it into R:
```{r include=FALSE}
SULFATE <- read.table("/home/edisz/Documents/Uni/Projects/blog/quantitative_ecotoxicology/data/p42.csv", 
                  header = TRUE, 
                  sep = ";")
```

```{r, eval=FALSE}
SULFATE <- read.table("p42.csv", 
                  header = TRUE, 
                  sep = ";")
```

Lets first have a look at the data via a violin plot:
```{r p43, message=FALSE}
require(ggplot2)
ggplot(SULFATE, aes(x = SITE, y = SO4)) +
  geom_violin()
```


It is quite easy to perform a wilcoxon-test with the function wilcox.test:
```{r}
wilcox.test(SO4 ~ SITE, data = SULFATE, correct = TRUE)
```
It works with the usual formula-notation, additional I specified the continuity correction.
For a one-sided test we can specify the argument 'alternative':
```{r}
wilcox.test(SO4 ~ SITE, data = SULFATE, correct = TRUE, alternative = "greater")
```

The p-values are the same as with SAS, however we get a warning since we have ties in our data (I don't know how SAS handles this):
> cannot compute exact p-value with ties

If we want to compute exact p-values in the presence of ties, we could use wilcox_test() from the coin package: 

```{r message=FALSE}
require(coin)
wilcox_test(SO4 ~ SITE, SULFATE, distribution="exact", conf.int = TRUE)
```
Here I also specified to output the 95%-Confidence-Interval via the 'conf.int argument.

Code and data are available at my [github-repo](https://github.com/EDiLD/r-ed/tree/master/quantitative_ecotoxicology) under file name 'p42'.
