
```{r setup, echo=FALSE, cache=FALSE}
options(scipen = 1, digits = 5)
```



### Quantitative Ecotoxicology, page 85, example 3.2, Nonlinear Least Squares

Get the data from [here](https://raw.github.com/EDiLD/r-ed/master/quantitative_ecotoxicology/data/p85.csv) and read it into R:

```{r include=FALSE}
OYSTERZN <- read.table("/home/edisz/Documents/Uni/Projects/blog/quantitative_ecotoxicology/data/p85.csv", 
                      header = TRUE, 
                      sep = ";")

```

```{r, eval=FALSE}
OYSTERZN <- read.table("p85.csv", 
                      header = TRUE, 
                      sep = ";")
OYSTERZN
```

```{r p85_raw}
plot(ZINC ~ DAY, data = OYSTERZN)
```

First we fit a **nonlinear Regression without weighting**.
```{r}
mod_noweight <- nls(ZINC ~ INITACT * exp((-(KE + 0.00283)) * DAY), 
           data = OYSTERZN,
           start = list(KE = 0.004, INITACT = 500))
```

So we fit the model

$$
 C_t = C_0 e^{-(k_{e1} + k_{e2}) t}
$$

to our data.

In the R formula **ZINC** corresponds to $C_t$, 
**INITACT** to $C_0$, 
**KE** to $k_{e1}$, 
**0.00283** is the decay rate constant for 65-Zn $k_{e2}$
and **DAY** to $t$.


Were are going to estimate **KE** and **INITACT** and also supplied some start-values for the algorithm.

We can look a the summary to get the estimates and standard error:
```{r}
summary(mod_noweight)
```

```{r include=FALSE}
sum_noweight <- summary(mod_noweight)
```

The resulting estimates of $k_e$ and $C_0$ are `r sum_noweight$coefficients[1, 1]` $\pm$ `r sum_noweight$coefficients[1, 2]` and  `r round(sum_noweight$coefficients[2, 1], 0)` $\pm$ `r round(sum_noweight$coefficients[2, 2], 0)`.

We can investige the residuals:

```{r p85_residuals_nls}
res_mod_noweight <- resid(mod_noweight)
plot(OYSTERZN$DAY, res_mod_noweight)
abline(h = 0)
```

Secondly, we run a **nonlinear regression with day-squared weighting**:

We use day^2 as weights and add there a column to our data:
```{r}
OYSTERZN$WNLIN <- OYSTERZN$DAY^2
```

We run again nls, but now we supply this new column as weights:
```{r}
mod_weight <- nls(ZINC ~ INITACT * exp((-(KE+0.00283))*DAY), 
                  data = OYSTERZN,
                  weights = OYSTERZN$WNLIN, 
                  start = list(KE = 0.004, INITACT = 500))
summary(mod_weight)
```

```{r include=FALSE}
sum_weight <- summary(mod_weight)
```


The estimates (`r sum_weight$coefficients[1, 1]` $\pm$ `r sum_weight$coefficients[1, 2]` and  `r round(sum_weight$coefficients[2, 1], 0)` $\pm$ `r round(sum_weight$coefficients[2, 2], 0)`) are quite similare to the nonweighted regression.

We could plot the two models and the data:

```{r p85_nls_fitted}
# extract the fitted values
fit_mod_noweight <- fitted(mod_noweight)
fit_mod_weight <- fitted(mod_weight)
# plot data
plot(ZINC ~ DAY, data = OYSTERZN)
# add fitted values
lines(OYSTERZN$DAY, fit_mod_noweight)
lines(OYSTERZN$DAY, fit_mod_weight, lty = "dashed")
# add legend
legend("topright", legend = c("nonweighted", "weighted"), lty=c("solid", "dashed"))
```


Finally we can also fit a **linear model** to the transformed Zinc-Concentrations:

First we ln-transform the concentrations:
```{r}
OYSTERZN$LZINC <- log(OYSTERZN$ZINC)
```

We see that the data is now linear:
```{r p85_raw2}
plot(LZINC ~ DAY, OYSTERZN)
```

And fit a linear regression:
```{r}
mod_lm <- lm(LZINC ~ DAY, data = OYSTERZN)
summary(mod_lm)
```

```{r include=FALSE}
sum_lm <- summary(mod_lm)
```

which is fitting the model
$$ln(Zn) = a * day + intercept$$, with a = `r round(sum_lm$coefficients[2, 1], 4)` and intercept = `r round(sum_lm$coefficients[1, 1], 2)`

Now plot data and model, as well as the residuals:
```{r p85_lm_fitted}
# fitted values
fit_mod_lm <- fitted(mod_lm)

# data + fitted
plot(LZINC ~ DAY, OYSTERZN)
lines(OYSTERZN$DAY, fit_mod_lm)

# residuals
plot(mod_lm, which = 1)
```


Code and data are available at my [github-repo](https://github.com/EDiLD/r-ed/tree/master/quantitative_ecotoxicology) under file name 'p85'.
