
```{r setup, echo=FALSE, cache=FALSE}
options(scipen = 1, digits = 5)
```



### Quantitative Ecotoxicology, page 85, example 3.2, Nonlinear Regression

Get the data from [here](https://raw.github.com/EDiLD/r-ed/master/quantitative_ecotoxicology/data/p85.csv) and read it into R:

```{r include=FALSE}
OYSTERZN <- read.table("/home/edisz/Documents/Uni/Projects/blog/quantitative_ecotoxicology/data/p85.csv", 
                      header = TRUE, 
                      sep = ";")

```
```{r, eval=FALSE}
OYSTERZN <- read.table("p85.csv", 
                      header = TRUE, 
                      sep = ";")
```
```{r}
head(OYSTERZN)
```
```{r p85_raw}
plot(ZINC ~ DAY, data = OYSTERZN)
```

First we fit a **nonlinear Regression without weighting**.
```{r}
mod_noweight <- nls(ZINC ~ INITACT * exp((-(KE + 0.00283)) * DAY), 
           data = OYSTERZN,
           start = list(KE = 0.004, INITACT = 500))
```

So we fit the model

![alt text](http://chart.apis.google.com/chart?cht=tx&chl=C_t%20=%20C_0%20e^{-%28k_{e1}%2Bk_{e2}%29%20t})

to our data.

In the R formula **ZINC** corresponds to ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=C_t) , 
**INITACT** to ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=C_0), 
**KE** to ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=k_{e1}), 
**0.00283** is the decay rate constant for 65-Zn ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=k_{e2})
and **DAY** to ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=t).


Were are going to estimate **KE** and **INITACT** and also supplied some start-values for the algorithm.

We can look a the summary to get the estimates and standard error:
```{r}
summary(mod_noweight)
```

```{r include=FALSE}
sum_noweight <- summary(mod_noweight)
```

The resulting estimates of ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=k_{e1}) and ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=C_0) are `r sum_noweight$coefficients[1, 1]` ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=\\pm) `r sum_noweight$coefficients[1, 2]` and  `r round(sum_noweight$coefficients[2, 1], 0)` ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=\\pm) `r round(sum_noweight$coefficients[2, 2], 0)`.

We can investigate the residuals, which show a clear pattern:

```{r p85_residuals_nls}
res_mod_noweight <- resid(mod_noweight)
plot(OYSTERZN$DAY, res_mod_noweight)
abline(h = 0)
```

Secondly, we run a **nonlinear regression with day-squared weighting**:

We use day^2 as weights and add there a column to our data:
```{r}
OYSTERZN$WNLIN <- OYSTERZN$DAY^2
```

We run again nls, but now we supply this new column as weights:
```{r}
mod_weight <- nls(ZINC ~ INITACT * exp((-(KE+0.00283))*DAY), 
                  data = OYSTERZN,
                  weights = OYSTERZN$WNLIN, 
                  start = list(KE = 0.004, INITACT = 500))
summary(mod_weight)
```

```{r include=FALSE}
sum_weight <- summary(mod_weight)
```


The estimates (`r sum_weight$coefficients[1, 1]` ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=\\pm) `r sum_weight$coefficients[1, 2]` and  `r round(sum_weight$coefficients[2, 1], 0)` ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=\\pm) `r round(sum_weight$coefficients[2, 2], 0)` are quite similar to the non weighted regression.

We could plot the two models and the data:

```{r p85_nls_fitted}
# extract the fitted values
fit_mod_noweight <- fitted(mod_noweight)
fit_mod_weight <- fitted(mod_weight)
# plot data
plot(ZINC ~ DAY, data = OYSTERZN)
# add fitted values
lines(OYSTERZN$DAY, fit_mod_noweight)
lines(OYSTERZN$DAY, fit_mod_weight, lty = "dashed")
# add legend
legend("topright", legend = c("nonweighted", "weighted"), lty=c("solid", "dashed"))
```


Finally we can also fit a **linear model** to the transformed Zinc-Concentrations:

First we ln-transform the concentrations:
```{r}
OYSTERZN$LZINC <- log(OYSTERZN$ZINC)
```

We see that the data has now linear trend:
```{r p85_raw2}
plot(LZINC ~ DAY, OYSTERZN)
```

And fit a linear regression:
```{r}
mod_lm <- lm(LZINC ~ DAY, data = OYSTERZN)
sum_lm <- summary(mod_lm)
sum_lm
```

which is fitting the model
![alt text](http://chart.apis.google.com/chart?cht=tx&chl=ln%28Zn%29%20=%20a%20*%20day%2Bintercept) with a = `r round(sum_lm$coefficients[2, 1], 4)` and intercept = `r round(sum_lm$coefficients[1, 1], 2)`

Now plot data and model, as well as the residuals:
```{r p85_lm_fitted}
# fitted values
fit_mod_lm <- fitted(mod_lm)

# data + fitted
plot(LZINC ~ DAY, OYSTERZN)
lines(OYSTERZN$DAY, fit_mod_lm)

# residuals
plot(mod_lm, which = 1)
```

The mean square error can be calculated from the summary:
```{r}
# MSE
sum_lm$sigma^2
```
From which we can get an unbiased estimate of $C_0$:
```{r}
# unbiased estimate for C_0: exp(MSE/2) * exp(Intercept)
exp(sum_lm$sigma^2 / 2) * exp(sum_lm$coefficients[1, 1])
```
where 
```{r eval=FALSE}
sum_lm$coefficients[1, 1]
```
extracts the intercept from the summary.

The estimated k in the summary output is `r sum_lm$coefficients[2, 1]` ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=\\pm) `r sum_lm$coefficients[2, 2]`, and ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=k_e%20=%20k%20-%20decayrate%20=%200.00531%20-%200.00283%20=%200.00248).

This result is similar to the weighted and non weighted nonlinear regression.
Again we have the same results as with SAS :) [Small deviations may be due to rounding error]




Code and data are available at my [github-repo](https://github.com/EDiLD/r-ed/tree/master/quantitative_ecotoxicology) under file name 'p85'.