
```{r setup, echo=FALSE, cache=FALSE}
options(scipen = 1, digits = 5)
```


### Quantitative Ecotoxicology, page 85, example 3.3, Backstripping

Get the data from [here](https://raw.github.com/EDiLD/r-ed/master/quantitative_ecotoxicology/data/p89.csv) and read it into R:

```{r eval=FALSE}
MERCURY <- read.table('p89.csv', 
                      header = TRUE, 
                      sep = ";")
```
```{r include=FALSE}
MERCURY <- read.table('/home/edisz/Documents/Uni/Projects/blog/quantitative_ecotoxicology/data/p89.csv', 
                      header = TRUE, 
                      sep = ";")
```
```{r}
head(MERCURY)
```
```{r p89_raw}
plot(PCI ~ DAY, data = MERCURY)
abline(v = 22)
```

We can identify two compartments on this plot: A slow one after day 22 and a fast one before day 22.

First we estimate ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=C_B) and ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=k_B) for the slow compartment, using linear regression of ln-transformed activity against day and predict from this slow compartment the activity over the whole period:

```{r}
# ln-transformation
MERCURY$LPCI <- log(MERCURY$PCI)
# fit linear model for day 31 to 94
mod_slow <- lm(LPCI ~ DAY, data = MERCURY[MERCURY$DAY > 22, ])
sum_mod_slow <- summary(mod_slow)
sum_mod_slow
exp(coef(mod_slow)[1])
```
So this gives us the model ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=C = 12468 e^{-0.0124 * Day}) for the slow component.

We do a bias correction and predict the activity for the whole data-range:
```{r}
# bias-correction
corr_mod_slow <- exp(sum_mod_slow$sigma^2 / 2) 
# add bias corrected predicts to data.frame
# predict takes the whole data.frame es newdata, so we get predicts for every day.
MERCURY$PRED <- exp(predict(mod_slow, newdata = MERCURY)) * corr_mod_slow
# save C_B and k_B as objects (used later...)
CB <- exp(coef(mod_slow)[1]) * corr_mod_slow
KCB <- abs(coef(mod_slow)[2])
```

The residuals from these predictions for day 3 to 22 are associated with the fast compartment.
And we fit a linear regression to the ln-transformed residuals for the fast compartment.

```{r p89_residuals}
plot(LPCI ~ DAY, data = MERCURY)
points(MERCURY$DAY[1:4], MERCURY$LPCI[1:4], pch = 16)
abline(a = log(CB), b = -KCB)
for(i in 1:4) {
  lines(c(MERCURY$DAY[i], MERCURY$DAY[i]), c(log(MERCURY$PRED[i]), MERCURY$LPCI[i]), lwd = 2)
}
```

```{r}
# extract residuals
MERCURY$ERROR  <- MERCURY$PCI - MERCURY$PRED
MERCURY$ERROR[1:4]
# fit linear model to ln(residuals) for Day 3 to 22
mod_fast <- lm(log(ERROR) ~ DAY, data = MERCURY[MERCURY$DAY < 22, ])
sum_mod_fast <- summary(mod_fast)
sum_mod_fast
exp(coef(mod_fast)[1])
```

So the model for the fast component is: ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=C = 36171 e^{-0.297 * Day})

```{r}
# bias correction
corr_mod_fast <- exp(sum_mod_fast$sigma^2 / 2) 
# save C_A and k_A as objects
CA <- exp(coef(mod_fast)[1]) * corr_mod_fast
KCA <- abs(coef(mod_fast)[2])
```


Now we have two models: one for the fast component and one for the slow component, and we can make a plot similar to Figure 8.1 in Newman and Clements (2008, pp. 119–120). 

```{r p89_backstripped}
plot(LPCI ~ DAY, data = MERCURY)
abline(mod_slow)
abline(mod_fast, lty = "dotted")
legend("topright", c("slow", "backstripped-fast"), lty=c("solid", "dashed"), cex = 0.8)
# Estimates
c(CA, KCA, CB, KCB)

```

We can use this estimates as start-values to fit a non-linear Model to the data (therefore we stored them into objects).

We want to fit the following model:
![alt text](http://chart.apis.google.com/chart?cht=tx&chl=C=C_A * e^{-k_A*Day}%2BC_B e^{-k_B*Day})

```{r}
nls_mod1 <- nls(PCI ~ CA * exp(-KCA * DAY) + CB * exp(-KCB * DAY), 
                data = MERCURY, 
                algorithm = "port",    # to use the bonds
                start = list(KCB = KCB, KCA = KCA, CB = CB, CA = CA),
                lower = c(0, 0, 5000, 20000), 
                upper = c(1, 1, 20000, 45000))
sum_nls_mod1 <- summary(nls_mod1)
sum_nls_mod1
```
* ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=C_A) is estimated as `r round(sum_nls_mod1$coefficients[4, 1], 0)` ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=\\pm) `r round(sum_nls_mod1$coefficients[4, 2], 0)`
* ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=C_B) is estimated as `r round(sum_nls_mod1$coefficients[3, 1], 0)` ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=\\pm) `r round(sum_nls_mod1$coefficients[3, 2], 0)`
* ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=k_A) is estimated as `r round(sum_nls_mod1$coefficients[2, 1], 3)` ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=\\pm) `r round(sum_nls_mod1$coefficients[2, 2], 3)`
* ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=k_A) is estimated as `r round(sum_nls_mod1$coefficients[1, 1], 3)` ![alt text](http://chart.apis.google.com/chart?cht=tx&chl=\\pm) `r round(sum_nls_mod1$coefficients[1, 2], 3)`


And finally we plot data and model.
```{r p89_nls}
plot(PCI ~ DAY, data = MERCURY, type = "n")
points(MERCURY$DAY, MERCURY$PCI, pch = ifelse(MERCURY$DAY <= 22, 16, 17))
# smooth line
pred_nls_mod1 <- predict(nls_mod1, newdata = data.frame(DAY = seq(0,100, 1)))
lines(seq(0,100, 1), pred_nls_mod1)
legend("topright", c("Fast", "slow"), pch=c(16,17))
```


Again we get nearly the same results with R, except for some differences in the linear models.

This is probably due to the bias-correction in slow-component-model.
We have a MSE of
```{r}
sum_mod_slow$sigma^2
```
which is identical to the book. From the previous example, the bias can be estimated as
![alt text](http://chart.apis.google.com/chart?cht=tx&chl=e^{MSE/2}):
```{r}
exp(sum_mod_slow$sigma^2 / 2)
```
which is different to the book (1.002).


I have no SAS at hand, so I cannot check this with SAS. However let me know if there is an error
in my calculations.


**Refs**

> Newman, Michael C., and William Henry Clements. Ecotoxicology: A Comprehensive Treatment. Boca Raton: Taylor /& Francis, 2008. 



Code and data are available at my [github-repo](https://github.com/EDiLD/r-ed/tree/master/quantitative_ecotoxicology) under file name 'p89'.