```{r setup, echo=FALSE, cache=FALSE}
options(scipen = 1, digits = 5)
```

Get the data from [here](https://raw.github.com/EDiLD/r-ed/master/quantitative_ecotoxicology/data/p94.csv) and read it into R:

```{r eval=FALSE}
LEAD <- read.table('p94.csv', 
                      header = TRUE, 
                      sep = ";")
```
```{r include=FALSE}
LEAD <- read.table('/home/edisz/Documents/Uni/Projects/blog/quantitative_ecotoxicology/data/p94.csv', 
                      header = TRUE, 
                      sep = ";")
```
```{r}
head(LEAD)
```


As always we first take a look at the data:
```{r p94_raw}
plot(LEAD ~ DAY, LEAD)
```

A simple power model may fit the data:

$$C_t = C_1~t^{−P}$$

We could fit such model as in example 3.3 via Nonlinear Least Squares or we could try to linearize the relationship by a ln-transform  of both DAY and LEAD:

```{r p94_linear}
LEAD$LLEAD <- log(LEAD$LEAD)
LEAD$LDAY <- log(LEAD$DAY)
plot(LLEAD ~ LDAY, LEAD)
```

Now we can us lm() to estimate the coefficients and check our model:

```{r}
# fit model
mod <- lm(LLEAD ~ LDAY, data = LEAD)
```

The residuals show no pattern:
```{r p94_residuals}
plot(mod, which = 1)
```

From the model-output:
```{r}
mod_sum <- summary(mod)
mod_sum
```


We see that out fitted model hast the formula:
$$Ln(LEAD) = 3.0008 - 0.272 ln(DAY)$$
with an R-squared of 0.77 and is statistically significant. The standard errors for the two parameters are 0.064 and 0.031.

So our backtransformed model would be:
$$ LEAD = exp(3.0008)~Day^{-0.272} = 20.68~Day^{-0.272}$$

Finally we can also plot our model:
```{r p94_model}
plot(LLEAD ~ LDAY, LEAD)
abline(mod)
```


Code and data are available at my [github-repo](https://github.com/EDiLD/r-ed/tree/master/quantitative_ecotoxicology) under filename 'p94'.

