In my first post I want to reproduce the results of:

***Van den Brink, P. J. & Braak, C. J. F. . Principal response curves: Analysis of time-dependent multivariate responses of biological community to stress. Environmental Toxicology and Chemistry 18, 138–148 (1999).***


Principle Response Curves (PRC) are commonly used for analyzing ecotoxicological mesocosm experiments. 
Here I will show how to run such an analysis with R.
I won't dig in the mathematics behind it, so this is up to the reader.


The data comes with the vegan package which we will also use for the PRC:

```{r message=FALSE}
require(vegan)
data(pyrifos)
head(pyrifos[ ,c(1:10)])
```

So rows are samplings and colums are the species (with abbreviated names), a normal species x sites matrix.
The colnames code treatment and time, but we must create these two factors as well as a factor coding the ditch:

```{r}
# time
week <- gl(11, 12, labels=c(-4, -1, 0.1, 1, 2, 4, 8, 12, 15, 19, 24))  
# treatment
dose <- factor(rep(c(0.1, 0, 0, 0.9, 0, 44, 6, 0.1, 44, 0.9, 0, 6), 11)) 
ditch <- gl(12, 1, length=132)
```


With this at hand we can calculate and plot the PRC using the prc function:
```{r prc1_plot1}
pyr_prc <- prc(response = pyrifos, treatment = dose, time = week)
plot(pyr_prc, scaling = 1)
```

This comes quite near to Figure 3 in the paper, except some scaling on the y-Axis and all the species scores are displayed.

Man, that was complicated...
In Figure 3 they ploted only the species with scores greater than 0.5 or less then -0.5.
We can access the species scores with the summary function:

```{r results='hide'}
sum_prc <- summary(pyr_prc)
# Species scores:
sum_prc$sp
```
```{r}
head(sum_prc$sp)
```

And then select only those species with greater or less scores:

```{r prc1_plot2}
plot(pyr_prc, select = abs(sum_prc$sp) > 0.5, scaling = 1)
```

OK, thats it for now. I think in following posts I will reproduced their tables and also show alternatives to the Principle Response Curves.











